Interpretable Neural Networks using EAGGA

### 2025-03-13 paper/main.tex
add HPO intro comments + formal definition of problem, start MOO bullets

\usepackage{mathtools}
\usepackage{bm}

@@

\subsection{Hyperparameter Optimization (HPO)}
- give intro to HPO = black-box optim problem -> we can apply any black box algo

@@

\subsection{Multi-Objective Optimization}
- intro MOO
\subsubsection{Pareto-optimality}
\subsubsection{A-priori}
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals

>>>>

\usepackage{mathtools}
\usepackage{amsmath}  % use argmin + argmax, cf. https://tex.stackexchange.com/a/5255
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{bm}

@@

\subsection{Hyperparameter Optimization (HPO)}
- in contrast to model parameters, which are optimised during training, hyperparameters are parameters describing the ML algo + are specified before training
- still often have great impact on final (trained) model performance -> can be optimised too
- formal definition
  * dataset $\mathcal{D}\subseteq\mathcal{X}\times\mathcal{Y}$ consisting of $n$ tuples $(\boldsymbol{x}^{(i)}, y^{(i)})\stackrel{i.i.d.}{\sim}\mathbb{P}_{\boldsymbol{x}y}$
    (data-generating distribution), i.e. $\mathcal{D}=\{(\boldsymbol{x}^{(i)}, y^{(i)})\}_{i=1}^n$
  * ML algo $\mathcal{I}(\cdot,\boldsymbol\lambda)$ with HP config $\boldsymbol\lambda\in\Lambda$ maps dataset to fitted model
    $f:\mathcal{X}\rightarrow\mathbb{R}^g\subseteq\mathcal{Y}$ (where $g\in\mathbb{N}$ is \# of classes of target),
    i.e. $\mathcal{I}:(\mathbb{D}\times\Lambda)\rightarrow\mathcal{H}$, with $\mathbb{D}$ space of all finite datasets and $\mathcal{H}$ hypothesis space (space of all models $f$)
  * denote $\mathcal{I}_{\boldsymbol\lambda}$ as (untrained) model with fixed HP config $\boldsymbol\lambda$
  * loss function $L$
  * goal of model training: for fixed $\boldsymbol\lambda$, have $\mathcal{I}$ find model $f$ minimising expected generalisation error
    $GE(\mathcal{I},\boldsymbol\lambda,\mathcal{D},L)=\mathbb{E}_{(\boldsymbol{x},y)\sim\mathbb{P}_{\boldsymbol{x}y}}[L(y,\mathcal{I}_{\boldsymbol\lambda}(\mathcal{D})(\boldsymbol{x}))]$
  * goal of HP optim: find $\boldsymbol\lambda$ minimising expected generalisation error, i.e. $\argmin_{\boldsymbol\lambda\in\Lambda} GE(\mathcal{I},\boldsymbol\lambda,\mathcal{D},L)$
- \citep[p. 3]{10.1145/3610536}
- HPO = black-box optim problem -> we can apply any black box algo to tune

@@

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
\subsubsection{Pareto-optimality}
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori
- simplest approach: different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\min_{x\in\mathcal{X}} w_i f_i(x)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - \citep[p. 11]{10.1145/3610536}
  * Chebyshev: provably converges to Pareto front
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals
